---
title: "NDVI Trends"
output: html_document
---

Analyis file fore runing trend tests on 614 watersheds to examine NDVI trends over time


## Load packages

```{r}
library(spdep)
library(ggplot2)
library(readxl)
library(ggplot2)
library(trend)
library(splm)
library(dplyr) 
```


## Example Code

```{r}
head(iris)
```
```{r}
iris %>%
   group_by(Species) %>%
   do(data.frame(., resid = residuals(lm(Sepal.Length ~ Sepal.Width, data=.))))
```

## Read in the data
```{r}
NDVI<-read_excel("X:/Projects/2N Software/Science/2N Journal Papers/Urban Greeness Tracking/NDVI Data/NDVI_apr8.xlsx")
View(NDVI)
```
## Plot time series P + T

```{r}
p <- ggplot(
  NDVI,
  aes(wat_yr, PPT, group = Site_No_ID, color = factor(EPAREGION))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Water Year", y = "P") +
  theme(legend.position = "top")
  (pegend.position = "top")
p
        
t <- ggplot(
  NDVI,
  aes(wat_yr, Temp, group = Site_No_ID, color = factor(EPAREGION))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Water Year", y = "T") +
  theme(legend.position = "top")
  (pegend.position = "top")
t
```


## Examine NDVI P + T dependence overall
```{r}
p<-ggplot(NDVI, aes(x=PPT, y=NDVI_urban)) + geom_point()
p
t<-ggplot(NDVI, aes(x=Temp, y=NDVI_urban, color= EPAREGION, alpha = 0.05)) + geom_point()
t
i<-ggplot(NDVI, aes(x=urban_imp_perc, y=NDVI_urban, color= EPAREGION, alpha = 0.05)) + geom_point()
i
```
## Examine site level correlations for P, T, I
```{r}
corT<-NDVI %>%
  group_by(Site_No_ID) %>%
  summarize(COR=cor(NDVI_urban,Temp))
corT
T<-as.data.frame(corT)

corP<-NDVI %>%
  group_by(Site_No_ID) %>%
  summarize(COR=cor(NDVI_urban,PPT))
corP
P<-as.data.frame(corP)

corI<-NDVI %>%
  group_by(Site_No_ID) %>%
  summarize(COR=cor(NDVI_urban,urban_imp_perc))
corI
I<-as.data.frame(corI)

t <-ggplot(corT, aes(x = COR)) + geom_histogram() + ggtitle("Temp")
t
p <-ggplot(corP, aes(x = COR)) + geom_histogram() + ggtitle("Precip")
p
i <-ggplot(corI, aes(x = COR)) + geom_histogram() + ggtitle("Impervious")
i
```

Correspondence with P and I is moderate to strong in most drainages, moderate with T. Conclusion - include both P + T in NDVI regressions

After performing trend tests on the rediduals, wondering if this correction is introducing additional error, particularly becuase we are applying it to all watersheds. If P + T don't have significant trends in time, perhaps we are better off using the raw NDVI

## Perform MK Trend tests on P + T 
```{r}

T.sites<-split(NDVI$Temp, forcats::fct_inorder(factor(NDVI$Site_No_ID)))

mk<- lapply(T.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out

sig<-ggplot(out, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ggtitle("T")

P.sites<-split(NDVI$PPT, forcats::fct_inorder(factor(NDVI$Site_No_ID)))

View(T.sites)

mk<- lapply(P.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out

sigP<-ggplot(out, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ggtitle ("P")
sigP

View(out)

```

Conclusion - more than half of the watersheds show trends over time in either P or T. This would indicatte that the NDVI should be corrected for these factors.

## Multivariate P+T NDVI Regressions + Store residuals
```{r}
NDVIr<-NDVI %>%
   group_by(Site_No_ID) %>%
   do(data.frame(., resid = residuals(lm(NDVI_urban ~ PPT + Temp, data=.))))

write.csv(NDVIr, file = "NDVIr.csv")

```
## Plot the NDVI residuals
```{r}
h <-ggplot(NDVIr, aes(x = resid)) + geom_histogram()
h
```



## Correlation plots for P + T residuals

```{r}
s <-ggplot(NDVIr, aes(x = wat_yr, y = resid)) + geom_point()
s
```


## Perform  MK Trend Tests + examine ouputs  P + T
```{r}

NDVIr.sites<-split(NDVIr$resid, forcats::fct_inorder(factor(NDVIr$Site_No_ID)))

mk<- lapply(NDVIr.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

outPT <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out

sigPT<-ggplot(outPT, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05)+ ylim(0,30) + ggtitle ("P+T")
sigPT

tau<-ggplot(out, aes(x= tau)) + geom_histogram() +  ggtitle ("P+T")
tau

```


# Put MK outputs back together with the sites data and extract significant trends P + T
```{r}
sites<-NDVIr %>%
select(Site_No_ID)
site<-unique(sites)

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
View(sigt)

msigt<- merge(sigt, NDVIr)
head(msigt)

psigt <-ggplot(msigt, aes(wat_yr,  resid, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile)") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt

```


## Multivariate T+P+I NDVI Regressions + store residuals in same data frame
```{r}
NDVIr2<-NDVI %>%
   group_by(Site_No_ID) %>%
   do(data.frame(., resid2 = residuals(lm(NDVI_urban ~ PPT + Temp + urban_imp_perc , data=.))))
head(NDVIr2)

```

## Correlation plots for P + T residuals

```{r}
s2 <-ggplot(NDVIr2, aes(x = wat_yr, y = resid2)) + geom_point()
s2

xx <-ggplot(NDVIr2, aes(x = Temp, y = PPT)) + geom_point()
> xx
>
```

```{r}
m<- lm(formula = resid2 ~ wat_yr, data = NDVIr2)
summary(m)
```

Conclusion - significant overall trend exits using a standared lm on residuals from P+T+I, but very weak. 
Next step = perform non-parametric MK test on P+T residuals for individual sites with I as covariate
OR use residuals from P+T+I regression - this is simpler


## Perform  MK Trend Tests + examine ouputs
```{r}

NDVIr2.sites<-split(NDVIr2$resid2, forcats::fct_inorder(factor(NDVIr2$Site_No_ID)))

mk<- lapply(NDVIr2.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

outPTI <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out

sigPTI<-ggplot(outPTI, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ggtitle ("P+T+I")
sigPTI

tau<-ggplot(out, aes(x= tau)) + geom_histogram() +  ggtitle ("P+T+I")
tau

```

Conclusion - 4 significant positive trends, 3 significant negative trends. 
Further analysis = try with I as direct covariat + perform change point analysis.


# Put MK outputs back together with the sites data and extract significant trends
```{r}
sites<-NDVIr2 %>%
select(Site_No_ID)
site<-unique(sites)

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
sigt

msigt<- merge(sigt, NDVIr2)
head(msigt)

psigt <-ggplot(msigt, aes(wat_yr,  resid2, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile)") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt

```
## Try the same trend tests on a shorter time period 1999-2019


```{r}
NDVIsh<- NDVI%>% filter (wat_yr > 1999)
summary(NDVIsh)
```


## Multivariate T+P+I NDVI Regressions + store residuals in same data frame for the short time series 2000-2019
```{r}
NDVIr3<-NDVIsh %>%
   group_by(Site_No_ID) %>%
   do(data.frame(., resid3 = residuals(lm(NDVI_urban ~ PPT + Temp + urban_imp_perc , data=.))))
NDVIr3
```

## Perform  MK Trend Tests + examine ouputs 2000-2019
```{r}
NDVIr3.sites<-split(NDVIr3$resid3, forcats::fct_inorder(factor(NDVIr3$Site_No_ID)))

mk<- lapply(NDVIr3.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out

sig<-ggplot(out, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ggtitle ("P+T+I")
sig
```

# Put MK outputs back together with the sites data and extract significant trends for 2000-2019

```{r}
sites<-NDVIr3 %>%
select(Site_No_ID)
site<-unique(sites)

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
sigt

msigt<-merge(sigt, NDVIr3)
head(msigt)


psigt <-ggplot(msigt, aes(wat_yr,  resid3, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt
```

Conclusion - fewer significant trends using the shorter time period. Positive trends Identified in two watersheds that were not identified in the longer time series

Perhaps we are introducing substantial error by using the regression residuals. Also try trend tests with no corrections for P + T with I as a partial correlation in the MK Test.



## Perform  MK Trend Tests + examine ouputs without correction for P + T 1985-2019
```{r}
NDVI.sites<-split(NDVI$NDVI_urban, forcats::fct_inorder(factor(NDVI$Site_No_ID)))
I.sites<-split(NDVI$urban_imp_perc, forcats::fct_inorder(factor(NDVI$Site_No_ID)))

NDVIurb <-as.data.frame(NDVI.sites)
I <-as.numeric(unlist(I.sites))

mk<- lapply(NDVI.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out3) <- NULL
out

sig<-ggplot(out, aes(x= p.value)) + geom_histogram()
sig
```

Try performing MK test on NDVI only corrected for I to see if results are any different

## Regression + store NDVI residuals for only I

```{r}
NDVIr4<-NDVI %>%
   group_by(Site_No_ID) %>%
   do(data.frame(., resid4 = residuals(lm(NDVI_urban ~urban_imp_perc , data=.))))
head(NDVIr4)
```

## Correlation plots for I residuals

```{r}
s <-ggplot(NDVIr4, aes(x = wat_yr, y = resid4)) + geom_point()
s
```

## Perform Partial MK Trend Tests for only I corrected residuals
```{r}
NDVIr4.sites<-split(NDVIr4$resid4, forcats::fct_inorder(factor(NDVIr4$Site_No_ID)))

mk<- lapply(NDVIr4.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out3) <- NULL
out

sig<-ggplot(out, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ggtitle ("I corr")
sig
```

## Put MK outputs back together with the sites data and extract significant trends for the I corrected NDVI

```{r}
sites<-NDVIr4 %>%
select(Site_No_ID)
site<-unique(sites)

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
sigt

msigt<-merge(sigt, NDVIr4)
head(msigt)


psigt <-ggplot(msigt, aes(wat_yr,  resid4, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt

```

Conclusions - no new trends are identified when omitting the P + T from the regression residuals, although the trend sloped changes slightly. Given that most sites have significant trends in time for either P or T, and most show moderate to strong association with NDVI , accounting for these variables seems warranted. 
Try performing MK test on NDVI only corrected for I to see if results are any different


## Regression + store NDVI residuals for only P + I corrected residuals

```{r}
NDVIr5<-NDVI %>%
   group_by(Site_No_ID) %>%
   do(data.frame(., resid5 = residuals(lm(NDVI_urban ~ PPT + urban_imp_perc , data=.))))
head(NDVIr5)
```

## Scatter plot for PPT + I corrected NDVI residuals

```{r}
s <-ggplot(NDVIr5, aes(x = wat_yr, y = resid5)) + geom_point()
s
```

## Perform Partial MK Trend Tests for only P + I corrected residuals
```{r}
NDVIr5.sites<-split(NDVIr5$resid5, forcats::fct_inorder(factor(NDVIr5$Site_No_ID)))

mk<- lapply(NDVIr5.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out

sig<-ggplot(out, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ggtitle ("I+P corr")
sig
```

## Put MK outputs back together with the sites data and extract significant trends for the P+I corrected NDVI

```{r}
sites<-NDVIr5 %>%
select(Site_No_ID)
site<-unique(sites)

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
sigt

msigt<-merge(sigt, NDVIr5)
head(msigt)


psigt <-ggplot(msigt, aes(wat_yr,  resid5, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile)") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt

```
Conclusions - only 1 site level trend change happens with the exclusio of T as a covariate. Some error is likely introduced at some sites where T correlation is not strong.


## Perform Partial MK Trend Tests for raw NDVI
```{r}
NDVI.sites<-split(NDVI$NDVI_urban, forcats::fct_inorder(factor(NDVI$Site_No_ID)))

mk<- lapply(NDVI.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

outRAW <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL

sigRAW<-ggplot(outRAW, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ylim(0, 30)+ ggtitle ("Raw NDVI")
sigRAW
```

## Plot all residual output p-values together
```{r}
outAll<-cbind(outRAW,outPT, outPTI)

names(outAll)[1]<-paste("p.valueRAW")
names(outAll)[6]<-paste("p.valuePT")
names(outAll)[11]<-paste("P.valuePTI") 

sigAll<-ggplot() + 
   geom_histogram(aes(x=outAll$p.valuePTI), color ="blueviolet", alpha =1) + xlim(0, 0.05) + ylim(0, 30)+
  geom_histogram(aes(x=outAll$p.valueRAW),color ="white", alpha = 1) + xlim(0, 0.05) + ylim(0, 30) +
  geom_histogram(aes(x=outAll$p.valuePT), color ="cyan", alpha = 1) + xlim(0, 0.05) + ylim(0, 30) 
 
sigAll
```


## Put MK outputs back together with the sites data and extract significant trends for the Raw NDVI

```{r}
sites<-NDVI%>%
select(Site_No_ID)
site<-unique(sites)

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
sigt
View(sigt)

msigt<-merge(sigt, NDVI)
head(msigt)
View(msigt)

library(ggforce)

psigt <-ggplot(msigt, aes(wat_yr,  NDVI_urban, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile)") +
  theme(legend.position = "top") + ylim(0.5, 1)

psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 1)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 2)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 3)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 4)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 5)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 6)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 7)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 8)


```

Raw NDVI trend tests show 307 significant trends, but many of these seem to be due to precipiatation or temperature patterns, which also show trends over time in hundreds of these watersheds.

## Apply the Partial MK test on NDVI corrected for P + T. Partial out the I term during the MK test

## Regression + store NDVI residuals for only P + I corrected residuals

```{r}
NDVIr6<-NDVI%>%
   group_by(Site_No_ID) %>%
   do(data.frame(., resid6 = residuals(lm(NDVI_urban ~ PPT, data=.))))
head(NDVIr6)
```

## Scatter plot for P + T corrected NDVI residuals

```{r}
s <-ggplot(NDVIr6, aes(x = wat_yr, y = resid6)) + geom_point()
s
```

## Perform Partial MK Trend Tests for only P + T corrected residuals, with I as covariate in trend test
```{r}
# split into columns
NDVIr6.sites<-split(NDVIr6$resid6, forcats::fct_inorder(factor(NDVIr6$Site_No_ID)))
I.sites<-split(NDVIr6$urban_imp_perc, forcats::fct_inorder(factor(NDVIr6$Site_No_ID)))

# change to numeric
r<-sapply(NDVIr6.sites, as.numeric)
i<-sapply(I.sites, as.numeric)

pmkt<- mapply(partial.mk.test, as.data.frame(r),as.data.frame(i))

library(reshape2)

d<-as.data.frame(pmkt)
df <- tibble::rownames_to_column(d, "VALUE")

dlong<-df%>%
    tidyr::pivot_longer(cols= starts_with("ID"), names_to="Site_No_ID", values_to = "result")
View(dlong)

pval<-subset(dlong, VALUE=="p.value")
pdata<-as.data.frame(pval)

p.value<-as.numeric(pdata[,3])->pdata[,3]
Site_No_ID<-as.character(pdata[,2])->pdata[2]

val<-data.frame(Site_No_ID, p.value)
View(val)


############# This this is the relevant output table from this analysis. Combined with Sen's Slope below.
pmk<- val %>% filter (p.value < 0.05)

View(pmk)

library(broom)
library(purrr)
library(tidyr)

sig<-ggplot(pmk, aes(x = p.value)) + geom_histogram() + xlim(0, 0.05) + ylim(0, 30) + ggtitle ("P+T, I PMK")
sig
```

## Plot significant trend sites of P+T residuals.
```{r}

msigt<-merge(pmk, NDVIr)

psigt <-ggplot(msigt, aes(wat_yr, resid, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt

psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 1)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 2)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 3)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 4)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 5)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 6)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 7)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 8)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 9)

```


## Plot significant trend sites of P+T+I residuals.
```{r}

msigt<-merge(pmk, NDVIr2)

psigt <-ggplot(msigt, aes(wat_yr, resid2, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt

psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 1)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 2)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 3)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 4)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 5)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 6)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 7)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 8)
psigt +facet_wrap_paginate(~Site_No_ID, nrow = 3, ncol = 3, page = 9)

```

Most of these slopes look pretty flat via the linear regression residuals P + T + I. The partial Mann-Kendall test must handle I as a covariate differntly than in the regression where it is also interacting with T + P

## Calculate Sen's Slope Estimator for the significant trends

```{r}

NDVIr.sites<-split(NDVIr$resid, forcats::fct_inorder(factor(NDVIr$Site_No_ID)))

s<- lapply(NDVIr.sites, sens.slope)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(s, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out


sig<-ggplot(out, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05)+ ylim(0, 30) + ggtitle ("P+T, I PMK")
sig

## Need to filter the sen plot by the significant trends from the PMK tests - output is 'pmk', but out does not have Site IDs attached



```

Since Sens's slop estimator has not had the impervious variable, it resluts in more p-values < 0.05 compared to the PMK test. This means that we can't just filter this output by the p-value, we need to join it with the PMK outputs


## Put MK outputs back together with the sites data and extract significant trends for the P + T, I

```{r}
NDVIr<-read_excel("X:/Projects/2N Software/Science/2N Journal Papers/Urban Greeness Tracking/R Urban Greenness/NDVIr.xlsx")


sites<-NDVIr %>%
select(Site_No_ID)
site<-unique(sites)

# Need to merge first on the PMK outputs - Sen's p-value is different. can't do this with no Site IDs in pmk

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
sigt

# This is th eoutput table with Sen's
msigt<-merge(sigt, NDVIr)
View(msigt)

# need to filter it for unique sites + combine with the PMK output to get slopes for the significant trends

sites<-msigt %>%
select(Site_No_ID,Sen.s.slope, statistic, Lat, Long, EPAREGION)
site<-unique(sites)


########### Output table with Sen's slope + PMK test p-values = 76 significant trends

mpmk<-merge(pmk,site)

sen<-ggplot(mpmk, aes(x= Sen.s.slope)) + geom_histogram() +  ylim(0, 12) + ggtitle ("P+T, I PMK")
sen

write.csv(mpmk, file = "mpmk.csv")

```

Changes are on the order of 0.002 NDVI/yr, which is about 0.07 NDVI for the entire study period.  
Most of these trends are negative

## Make some maps from the trend results

```{r}
library(maps)
library(ggthemes)
library(mapproj)
library(mapdata)

## These maps work, but only data without a usa polygon

usaMap <- map_data("usa")
MainStates<-map_data("state", color="white")

# Data for usa polygon
usa <- ggplot() +
  borders("state", colour = "white", fill = "gray92") + coord_fixed(1.3) +
  theme_void()
usa

# Try another way to draw USA polygon
usa2<- ggplot() + 
  geom_polygon(data=MainStates, aes(x=long, y=lat, group=group),
  color=NA, fill="grey95", states = TRUE )  + coord_fixed(1.3) + theme_void()
usa2

cities<-data(world.cities, labesl = TRUE)

MainCities <- filter(us.cities, long>=-130)
BigCities <-filter(MainCities, pop > 500000)


clab <- usa2 + geom_point(data=BigCities, aes(x=long, y=lat), size=1, alpha = 1) + geom_text(aes(x=long, y= lat, label = name),
        data= BigCities,
        alpha = 1,
        color = "black",
        size = 3)

usac <- usa2 + geom_point(data=BigCities, aes(x=long, y=lat), size=1, alpha = 1)
usac

```


```{r}
# NDVI Trends map

map <- usa +
  geom_point(aes(x = Long, y = Lat, color = Sen.s.slope, size = p.value),
          data = mpmk, 
          alpha = .4) + scale_size_continuous(range = c(1, 10)) +
          labs(size = 'p.value') + theme_void() 
map
  

# Map of initial NDVI
m <- map + scale_color_gradient(low = "mediumpurple4", high = "springgreen2") 
m

nap <- usa +
  geom_point(aes(x = Long, y = Lat, color = NDVI_urban, size = Urban_area_km2),
          data = NDVIr, 
          alpha = .4) + scale_size_continuous(range = c(1, 10)) +
          labs(size = 'Urban_area_km2') + theme_void() 
nap
  

# colors + legends
n <- nap + scale_color_gradient(low = "mediumpurple4", high = "springgreen2") 
n




```




## Multivariate T+P+I NDVI Regressions + store P + T  residuals in same data frame for the short time series 2000-2019

```{r}
NDVIr7<-NDVIsh %>%
   group_by(Site_No_ID) %>%
   do(data.frame(., resid3 = residuals(lm(NDVI_urban ~ PPT + Temp, data=.))))
NDVIr7
```

## Perform  MK Trend Tests + examine ouputs 2000-2019
```{r}
NDVIr7.sites<-split(NDVIr3$resid7, forcats::fct_inorder(factor(NDVIr7$Site_No_ID)))

mk<- lapply(NDVIr7.sites, mk.test)

library(broom)
library(purrr)
library(tidyr)

out <- do.call(rbind, lapply(mk, function(x) 
  cbind(data.frame(p.value =  x$p.value, statistic  = x$statistic), 
          data.frame(as.list(x$estimates)))))
row.names(out) <- NULL
out

sig<-ggplot(out, aes(x= p.value)) + geom_histogram() + xlim(0, 0.05) + ggtitle ("P+T 2000-2019")
sig
```

# Put MK outputs back together with the sites data and extract significant trends for 2000-2019

```{r}


sites<-NDVIr3 %>%
select(Site_No_ID)
site<-unique(sites)

siten <- tibble::rowid_to_column(site, "ID")
outn <- tibble::rowid_to_column(out, "ID")

siteout<-merge(siten,outn)

# Extract sites with significant trends and plot results 

sigt<- siteout %>% filter (p.value < 0.05)
sigt

msigt<-merge(sigt, NDVIr7)
head(msigt)


psigt <-ggplot(msigt, aes(wat_yr,  resid7, group = Site_No_ID)) + 
  geom_point()  + geom_smooth(method = "lm") + scale_color_viridis_d() + 
  labs(x = "Water Year", y = "NDVI (90th percentile") +
  theme(legend.position = "top") + facet_wrap (~Site_No_ID)
psigt



